---
title: "FAMD"
author: "Alayna Johnson"
date: "2024-09-19"
output: html_document
---


```{r}
library(tidyverse)
library(FactoMineR)

ctrl_vars <- read_csv("ctrl_vars.csv")
```

```{r}
ctrl_vars <- ctrl_vars %>%
  mutate(white_perc = (white_total/total_population)) %>%
  mutate(blkAA_perc = (blkAA_total/total_population)) %>%
  mutate(AIAN_perc = (AIAN_total/total_population)) %>%
  mutate(asian_perc = (asian_total/total_population)) %>%
  mutate(PI_perc = (PI_total/total_population)) %>%
  mutate(other_perc = (other_total/total_population)) %>%
  mutate(twomore_perc = (twomore_total/total_population)) %>%
  mutate(foreign_perc = (foreign_born/population_2022)) %>%
  mutate(able_to_work = (population_2022/total_population)) %>%
  select(-white_total, -blkAA_total,-AIAN_total, -asian_total, -PI_total, -other_total, -twomore_total, -total_population, -foreign_born, -population_2022)



```


```{r}

clean_vars <- ctrl_vars %>%
  mutate(
    leg_ctrl = as.factor(leg_ctrl),
    gov_party = as.factor(gov_party),
    state_ctrl = as.factor(state_ctrl),
    det_sentencing = as.factor(det_sentencing)
  )
```




```{r}
small_data <- clean_vars %>%
  select(foreign_perc, poverty_perc, no_highschool, unemployment_perc, state_ctrl, bills_enacted, reincarceration)


# write.csv(small_data, "small_control.csv", row.names = FALSE)

small_analysis <- FAMD(small_data, ncp = 3)

summary(small_analysis)
```

```{r}
factors <- small_analysis$ind$coord

```


```{r}
pc_data <- as.data.frame(factors)

regression <- read.csv("regression.csv")

final_df_orig <- cbind(pc_data, regression)

final_df <- final_df_orig %>%
  select(-state) %>%
  mutate(scaled_fair_chance = fair_chance/100) %>%
  select(-fair_chance)
```


```{r}
library(betareg)

model <- lm(scaled_fair_chance ~., data = final_df)
summary(model)

```

phi: The precision coefficient is highly significant (p = 6.82e-07) and very large, meaning that there is low dispersion (i.e., the dependent variable is tightly clustered around the mean).

```{r}
final_df %>%
  lm(fair_chance ~ ) %>%
  summary()

```


```{r}
library(ggplot2)
library(broom)

# Extract the coefficients and confidence intervals
coef_data <- tidy(model, conf.int = TRUE)

# Plot the coefficients
ggplot(coef_data, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange() +
  coord_flip() +  # Flip coordinates for better readability
  labs(title = "Beta Regression Coefficients", 
       x = "Predictor", 
       y = "Estimate") +
  theme_minimal()

```

```{r}
# Predict the fitted values
final_df$fitted_values <- predict(model, type = "response")

# Plot actual vs fitted values
ggplot(final_df, aes(x = fitted_values, y = scaled_fair_chance)) +
  geom_point(alpha = 0.6) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "firebrick") + 
  labs(title = "Fitted vs Actual Values",
       x = "Fitted Values",
       y = "Actual Values") +
  theme_minimal()

```

Suggests that the model is performing well at predicting scaled_fair_chance because the points stay close to the line (perfect predictions)

```{r}
# Extract residuals
final_df$residuals <- residuals(model, type = "quantile")

# Plot residuals
ggplot(final_df, aes(x = fitted_values, y = residuals)) +
  geom_point(alpha = 0.6) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Fitted Values",
       x = "Fitted Values",
       y = "Residuals") +
  theme_minimal()

```

```{r}
# Calculate the precision (phi)
phi_value <- 1 / predict(model, type = "precision")

# Plot the precision parameter against fitted values
ggplot(final_df, aes(x = fitted_values, y = phi_value)) +
  geom_point(alpha = 0.6) +
  labs(title = "Precision vs Fitted Values",
       x = "Fitted Values",
       y = "Precision (Phi)") +
  theme_minimal()

```


Are there any transformations or interactions that might better capture the relationship between the predictors and the response??

```{r}
model_neg <- betareg(scaled_fair_chance ~neg_combo_index, data = final_df)
summary(model_neg)

model_pos <- betareg(scaled_fair_chance ~tot_pos_index, data = final_df)
summary(model_pos)

```

```{r}
model_less <- betareg(scaled_fair_chance ~tot_pos_index+neg_combo_index+Dim.1+Dim.2+Dim.3,data = final_df)
summary(model_less)
```
**Removing California**

```{r}
no_cali_data <- clean_vars %>%
  filter(state !="California") %>%
  select(foreign_perc, poverty_perc, no_highschool, unemployment_perc, state_ctrl, bills_enacted, reincarceration)


no_cali_analysis <- FAMD(no_cali_data, ncp = 3)

```

```{r}
factors2 <- no_cali_analysis$ind$coord
```


```{r}
pc_data2 <- as.data.frame(factors2)

regression <- read.csv("regression.csv")

regression <- regression %>%
  filter(state != "California")

no_cali_df <- cbind(pc_data2, regression)

final_df <- no_cali_df %>%
  select(-state) %>%
  mutate(scaled_fair_chance = fair_chance/1000) %>%
  select(-fair_chance)
```



```{r}
model <- betareg(scaled_fair_chance ~., data = final_df)
summary(model)
```

