---
title: "Association between different positive policy and fair chance job posting rates"
author: "Zofia C. Stanley"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=TRUE}
library(ggplot2)

# Set parameter to overwrite plots
overwrite_plots = FALSE

# Read in the data
regression_data <- read.csv("regression.csv")
employment_data <- read.csv("../Employment/Indeed/indeed-retail-foodservice-fairchance.csv")
control_data <- read.csv("control_indices.csv")
pos_pol_data <- read.csv("../Policy/PositiveIndex.csv")
pos_pol_data <- pos_pol_data[1:50, ]
```

# Introduction
This analysis aims to explore the relationship between different positive policies, and fair chance hiring outcomes in retail and food service industries. We use scatter plots, linear regression models, and chi-square tests to understand these relationships.

# Scatter Plot: Positive Policies vs Collateral Consequences
We begin by visualizing the relationship between positive policies and collateral consequences.

```{r scatter-pos-neg, echo=TRUE}
# Create a scatter plot of positive policy vs negative policy
plot_pos_vs_neg <- ggplot(regression_data, aes(x = neg_combo_index, y = tot_pos_index)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Scatter Plot of Positive Policies vs Collateral Consequences",
    x = "Weighted Number of Collateral Consequences",
    y = "Positive Policy Index"
  ) +
  xlim(0, 900) +
  ylim(0, 1) +
  theme_minimal()

if (overwrite_plots) {
  pdf("scatter_pos_neg.pdf")
  print(plot_pos_vs_neg)
  dev.off()
}

print(plot_pos_vs_neg)
```

**Interpretation**: The scatter plot shows the relationship between the weighted number of collateral consequences and the positive policy index. Visually, we see a strong positive association between the weighted number of collateral consequences and positive policy index. 

# Scatter Plots: Ban the Box vs Policy Indices
We next look at how fair chance job posting rates relate to ban the box positive index.

```{r scatter-pos-fairchance, echo=TRUE}
# Create a scatter plot of positive policy index vs fair chance job posting rate (retail + food service)
plot_btb_vs_fairchance_rate <- ggplot(mapping=aes(y = employment_data$fair_chance_rate, x = pos_pol_data$Ban.the.Box.Index)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Scatter Plot of Fair Chance Job Posting Rate vs Ban the Box Index",
    x = "Ban the Box Index",
    y = "Fair Chance Job Posting Rate in Retail + Food Service"
  ) +
  theme_minimal()

if (overwrite_plots) {
  pdf("scatter_btb_fairchance_industry.pdf")
  print(plot_btb_vs_fairchance_rate)
  dev.off()
}
print(plot_btb_vs_fairchance_rate)
```

```{r scatter-neg-fairchance, echo=TRUE}
# Create a scatter plot of record clearing index vs fair chance job posting rate in retail and food service
plot_rec_clear_vs_fairchance_rate <- ggplot(mapping = aes(y = employment_data$fair_chance_rate, x = pos_pol_data$Record.Clearing.Index)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Scatter Plot of Fair Chance Job Posting Rate vs Record Clearing Index",
    x = "Record Clearing Index",
    y = "Fair Chance Job Posting Rate in Retail + Food Service"
  ) +
  theme_minimal()

if (overwrite_plots) {
  pdf("scatter_neg_fairchance_industry.pdf")
  print(plot_rec_clear_vs_fairchance_rate)
  dev.off()
}
print(plot_rec_clear_vs_fairchance_rate)
```

# Linear Models and Residual Analysis
To control for macroeconomic, socioeconomic, and political factors, we ran several linear models and used residuals to further explore relationships.

```{r linear-models, echo=TRUE}
# Run a linear model for fair_chance_rate vs the three columns of control_data
lm_fairchance_controls <- lm(employment_data$fair_chance_rate ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity)

# Save the residuals of the model to employment_data$fair_chance_rate_residuals
employment_data$fair_chance_rate_residuals <- residuals(lm_fairchance_controls)

# Print the summary of the linear model
summary(lm_fairchance_controls)

# Run a linear model for btb_index vs the three columns of control_data
lm_model_btb <- lm(pos_pol_data$Ban.the.Box.Index ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity)

# Save the residuals of the model to regression_data$tot_pos_index_residuals
regression_data$btb_index_residuals <- residuals(lm_model_btb)

# Print the summary of the linear model
summary(lm_model_btb)

# Run a linear model for recordclearing vs the three columns of control_data
lm_model_rec_clear <- lm(pos_pol_data$Record.Clearing.Index ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics  + control_data$diversity)

# Save the residuals of the model to regression_data$neg_combo_index_residuals
regression_data$rec_clear_index_residuals <- residuals(lm_model_rec_clear)

# Print the summary of the linear model
summary(lm_model_rec_clear)
```

# Residualized Analysis
We used residualized values to examine relationships after controlling for external factors.

```{r scatter-residual-pos-fairchance, echo=TRUE}
# Scatter plot of residualized positive policy vs. residualized fair chance job posting rate
plot_residual_btb_vs_fairchance <- ggplot(data = data.frame(y = employment_data$fair_chance_rate_residuals, x = regression_data$btb_index_residuals)) +
  geom_point(aes(x = x, y = y)) +
  geom_smooth(aes(x = x, y = y), method="lm", se=FALSE)+
  labs(
    title = "Scatter Plot of Residualized Ban the Box vs. Residualized Fair Chance Job Posting Rate",
    y = "Residualized Fair Chance Job Posting Rate, Retail + Food Service",
    x = "Residualized Ban the Box Index"
  ) +
  theme_minimal()
if (overwrite_plots) {
  pdf("scatter_residual_btb_fairchance.pdf")
  print(plot_residual_btb_vs_fairchance)
  dev.off()
}
print(plot_residual_btb_vs_fairchance)
```

```{r scatter-residual-neg-fairchance, echo=TRUE}
# Scatter plot of residualized negative policy vs. residualized fair chance job posting rate
plot_residual_rec_clear_vs_fairchance <- ggplot(data = data.frame(y = employment_data$fair_chance_rate_residuals, x = regression_data$rec_clear_index_residuals)) +
  geom_point(aes(x = x, y = y)) +
  geom_smooth(aes(x = x, y = y), method="lm", se=FALSE)+
  labs(
    title = "Scatter Plot of Residualized Record Clearing Index vs. Residualized Fair Chance Job Posting Rate",
    y = "Residualized Fair Chance Job Posting Rate, Retail + Food Service",
    x = "Residualized Record Clearing Index"
  ) +
  theme_minimal()
if (overwrite_plots) {
  pdf("scatter_residual_rec_clear_fairchance.pdf")
  print(plot_residual_rec_clear_vs_fairchance)
  dev.off()
}
print(plot_residual_rec_clear_vs_fairchance)
```


```{r additional-linear-models, echo=TRUE}
# Run a linear model for fair_chance_rate vs tot_pos_index, neg_combo_index, and control_data
lm_fair_chance_rate_vs_all_factors <- lm(employment_data$fair_chance_rate ~ regression_data$tot_pos_index + regression_data$neg_combo_index + control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity)

# Print the summary of the linear model
summary(lm_fair_chance_rate_vs_all_factors)

# Run a linear model for residualized fair_chance_rate vs residualized tot_pos_index and neg_combo_index
lm_residual_fair_chance_rate_vs_residual_policies <- lm(employment_data$fair_chance_rate_residuals ~ regression_data$tot_pos_index_residuals + regression_data$neg_combo_index_residuals)

# Print the summary of the linear model
summary(lm_residual_fair_chance_rate_vs_residual_policies)

# Run a linear model for residualized tot_pos_index vs residualized neg_combo_index
lm_residual_positive_policy_vs_residual_negative_combo <- lm(regression_data$tot_pos_index_residuals ~ regression_data$neg_combo_index_residuals)

# Print the summary of the linear model
summary(lm_residual_positive_policy_vs_residual_negative_combo)
```


# Chi-Square Test of Independence
Finally, we perform chi-square tests to understand if having an above-average policy index is associated with above-average employment outcomes.

```{r chi-square-test-pos, echo=TRUE}
# Create binary variables for "above or below average" for each measure (Positive Policy Index)
policy_above_avg_pos <- ifelse(regression_data$btb_index_residuals > 0, "Above Avg", "Below Avg")
employment_above_avg <- ifelse(employment_data$fair_chance_rate_residuals > 0, "Above Avg", "Below Avg")

# Create a contingency table based on the two categorical variables (Positive Policy Index)
contingency_table_pos <- table(policy_above_avg_pos, employment_above_avg)

# Print the contingency table (Positive Policy Index)
print(contingency_table_pos)

# Run the chi-square test of independence (Positive Policy Index)
chi_squared_result_pos <- chisq.test(contingency_table_pos)

# Print the test result (Positive Policy Index)
print(chi_squared_result_pos)
```
**Interpretation:** The chi-square test of independence for the ban the box has a very high p value demonstrating that an above average ban the box policy index is not statistically significantly correlated with an above average fair chance hiring.

```{r chi-square-test-neg, echo=TRUE}
# Create binary variables for "above or below average" for each measure (Record Clearing Index)
policy_above_avg_rec_clear <- ifelse(regression_data$rec_clear_index_residuals > 0, "Above Avg", "Below Avg")
employment_above_avg <- ifelse(employment_data$fair_chance_rate_residuals > 0, "Above Avg", "Below Avg")

# Create a contingency table based on the two categorical variables (Positive Policy Index)
contingency_table_rec_clear <- table(policy_above_avg_rec_clear, employment_above_avg)

# Print the contingency table (Positive Policy Index)
print(contingency_table_rec_clear)

# Run the chi-square test of independence (Positive Policy Index)
chi_squared_result_rec_clear <- chisq.test(contingency_table_rec_clear)

# Print the test result (Positive Policy Index)
print(chi_squared_result_rec_clear)
```

**Interpretation:** The chi-square test of independence for the record clearing index has a p value of 0.4302 meaning there is no correlation between having an above average record clearing index and above average engagement in fair chance hiring.

**Residuals and Modeling for Licencing and Credentialing**

```{r}
# Run a linear model for neg_combo_index vs the three columns of control_data
lm_model_landc <- lm(pos_pol_data$Licensing.and.Credentialing.Index ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity)

# Save the residuals of the model to regression_data$neg_combo_index_residuals
regression_data$landc_index_residuals <- residuals(lm_model_landc)

# Print the summary of the linear model
summary(lm_model_landc)

# Scatter plot of residualized licensing and credentialiing vs. residualized fair chance job posting rate
plot_residual_landc_vs_fairchance <- ggplot(data = data.frame(y = employment_data$fair_chance_rate_residuals, x = regression_data$landc_index_residuals)) +
  geom_point(aes(x = x, y = y)) +
  geom_smooth(aes(x = x, y = y), method="lm", se=FALSE)+
  labs(
    title = "Scatter Plot of Residualized Licensing and Credentialing Index vs. Residualized Fair Chance Job Posting Rate",
    y = "Residualized Fair Chance Job Posting Rate, Retail + Food Service",
    x = "Residualized Licensing and Credentialing Index"
  ) +
  theme_minimal()
if (overwrite_plots) {
  pdf("scatter_residual_landc_r_fairchance.pdf")
  print(plot_residual_landc_vs_fairchance)
  dev.off()
}
print(plot_residual_landc_vs_fairchance)


```

**Chi-Squared Tests** For Licencing and Credentialing

```{r}


# Create binary variables for "above or below average" for each measure (Licencing and Credentialing)
policy_above_avg_landc <- ifelse(regression_data$landc_index_residuals > 0, "Above Avg", "Below Avg")
employment_above_avg <- ifelse(employment_data$fair_chance_rate_residuals > 0, "Above Avg", "Below Avg")

# Create a contingency table based on the two categorical variables (Positive Policy Index)
contingency_table_landc <- table(policy_above_avg_landc, employment_above_avg)

# Print the contingency table (Positive Policy Index)
print(contingency_table_landc)

# Run the chi-square test of independence (Positive Policy Index)
chi_squared_result_landc <- chisq.test(contingency_table_landc)

# Print the test result (Positive Policy Index)
print(chi_squared_result_landc)


```

**Residuals and Modeling for Certificates of Rehabilitation**

```{r}
# Run a linear model for neg_combo_index vs the three columns of control_data
lm_model_rehab <- lm(pos_pol_data$Certificate.of.Rehabilitation.Index ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics  + control_data$diversity)

# Save the residuals of the model to regression_data$neg_combo_index_residuals
regression_data$rehab_index_residuals <- residuals(lm_model_rehab)

# Print the summary of the linear model
summary(lm_model_rehab)

# Scatter plot of residualized licensing and credentialiing vs. residualized fair chance job posting rate
plot_residual_rehab_vs_fairchance <- ggplot(data = data.frame(y = employment_data$fair_chance_rate_residuals, x = regression_data$rehab_index_residuals)) +
  geom_point(aes(x = x, y = y)) +
  geom_smooth(aes(x = x, y = y), method="lm", se=FALSE)+
  labs(
    title = "Scatter Plot of Residualized Cert. of Rehabilitation vs. Residualized Fair Chance Job Posting Rate",
    y = "Residualized Fair Chance Job Posting Rate, Retail + Food Service",
    x = "Residualized Cert. of Rehabilitation Index"
  ) +
  theme_minimal()
if (overwrite_plots) {
  pdf("scatter_residual_rehab_r_fairchance.pdf")
  print(plot_residual_rehab_vs_fairchance)
  dev.off()
}
print(plot_residual_rehab_vs_fairchance)


```

**Chi-Squared Tests** For Licencing and Credentialing

```{r}


# Create binary variables for "above or below average" for each measure (Licencing and Credentialing)
policy_above_avg_rehab <- ifelse(regression_data$rehab_index_residuals > 0, "Above Avg", "Below Avg")
employment_above_avg <- ifelse(employment_data$fair_chance_rate_residuals > 0, "Above Avg", "Below Avg")

# Create a contingency table based on the two categorical variables (Positive Policy Index)
contingency_table_rehab <- table(policy_above_avg_rehab, employment_above_avg)

# Print the contingency table (Positive Policy Index)
print(contingency_table_rehab)

# Run the chi-square test of independence (Positive Policy Index)
chi_squared_result_rehab <- chisq.test(contingency_table_rehab)

# Print the test result (Positive Policy Index)
print(chi_squared_result_rehab)


```
**Interpretation** NO SIGNIFICANCE WHATSOEVER


**Residuals and Modeling for Liability** 

```{r}
# Run a linear model for neg_combo_index vs the three columns of control_data
lm_model_liability <- lm(pos_pol_data$Liability.Index ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity)

# Save the residuals of the model to regression_data$neg_combo_index_residuals
regression_data$liability_index_residuals <- residuals(lm_model_liability)

# Print the summary of the linear model
summary(lm_model_liability)

# Scatter plot of residualized licensing and credentialiing vs. residualized fair chance job posting rate
plot_residual_liability_vs_fairchance <- ggplot(data = data.frame(y = employment_data$fair_chance_rate_residuals, x = regression_data$liability_index_residuals)) +
  geom_point(aes(x = x, y = y)) +
  geom_smooth(aes(x = x, y = y), method="lm", se=FALSE)+
  labs(
    title = "Scatter Plot of Residualized Liablity vs. Residualized Fair Chance Job Posting Rate",
    y = "Residualized Fair Chance Job Posting Rate, Retail + Food Service",
    x = "Residualized Liability Index"
  ) +
  theme_minimal()
if (overwrite_plots) {
  pdf("scatter_residual_liability_fairchance.pdf")
  print(plot_residual_liability_vs_fairchance)
  dev.off()
}
print(plot_residual_liability_vs_fairchance)

```
```{r}


# Create binary variables for "above or below average" for each measure (Liability Index)
policy_above_avg_liability <- ifelse(regression_data$liability_index_residuals > 0, "Above Avg", "Below Avg")
employment_above_avg <- ifelse(employment_data$fair_chance_rate_residuals > 0, "Above Avg", "Below Avg")

# Create a contingency table based on the two categorical variables (LiabilityIndex)
contingency_table_liablility <- table(policy_above_avg_liability, employment_above_avg)

# Print the contingency table (Liability Index)
print(contingency_table_liablility)

# Run the chi-square test of independence (Liability Index)
chi_squared_result_liablility <- chisq.test(contingency_table_liablility)

# Print the test result (Liability Index)
print(chi_squared_result_liablility)


```
**Interpretation** No significance 