---
title: "CleanCode"
author: "Jocelyn Bliven"
date: "2024-11-05"
output: pdf_document
---

```{r setup, include=TRUE}

library(ggplot2)

# Set parameter to overwrite plots
overwrite_plots = FALSE

# Read in data: get 
pol_indices <- read.csv("regression.csv")
employment_data <- read.csv("wide_complete_industry.csv")
control_data <- read.csv("control_indices.csv")
employ_z_scores <- read.csv("employment_z_scores.csv")
indiv_pol_data <- read.csv("../Policy/PositiveIndex.csv")
indiv_pol_data <- indiv_pol_data[1:50, ]

```

# Linear Models and Residual Analysis of RETAIL
To control for macroeconomic, socioeconomic, and political factors, we ran several linear models and used residuals to further explore relationships.

```{r linear-models, echo=TRUE}
# Run a linear model for fair_chance_rate vs the 5 types of control_data
lm_fairchance_controls <- lm(employment_data$avg_retail ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity + employ_z_scores$rtrade_employment_z)

# Save the residuals of the model to employment_data$fair_chance_rate_residuals
employment_data$retail_residuals <- residuals(lm_fairchance_controls)

# Print the summary of the linear model
summary(lm_fairchance_controls)

# Run a linear model for the total positive index vs the 5 types of control_data
lm_model_pos_index <- lm(pol_indices$tot_pos_index ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity + employ_z_scores$rtrade_employment_z)

# Save the residuals of the model to regression_data$tot_pos_index_residuals
pol_indices$pos_index_residuals <- residuals(lm_model_pos_index)

# Print the summary of the linear model
summary(lm_model_pos_index)

```
# Residualized Analysis
We used residualized values to examine relationships after controlling for external factors.

```{r scatter-residual-pos-fairchance, echo=TRUE}
# Scatter plot of residualized positive policy vs. residualized fair chance job posting rate
plot_residual_pos_vs_fairchance <- ggplot(data = data.frame(y = employment_data$retail_residuals, x = pol_indices$pos_index_residuals)) +
  geom_point(aes(x = x, y = y)) +
  geom_smooth(aes(x = x, y = y), method="lm", se=FALSE)+
  labs(
    title = "Scatter Plot of Residualized Positive Policy vs. Residualized Fair Chance Job Posting Rate",
    y = "Residualized Fair Chance Job Posting Rate, Retail + Food Service",
    x = "Residualized Postitve Index"
  ) +
  theme_minimal()
if (overwrite_plots) {
  pdf("plot_residual_pos_vs_fairchance.pdf")
  print(plot_residual_pos_vs_fairchance)
  dev.off()
}
print(plot_residual_pos_vs_fairchance)
```

# Chi-Square Test of Independence
Finally, we perform chi-square tests to understand if having an above-average policy index is associated with above-average employment outcomes.

```{r chi-square-test-pos, echo=TRUE}
# Create binary variables for "above or below average" for each measure (Positive Policy Index)
policy_above_avg_pos <- ifelse(pol_indices$pos_index_residuals > 0, "Above Avg", "Below Avg")
employment_above_avg <- ifelse(employment_data$retail_residuals > 0, "Above Avg", "Below Avg")

# Create a contingency table based on the two categorical variables (Positive Policy Index)
contingency_table_pos <- table(policy_above_avg_pos, employment_above_avg)

# Print the contingency table (Positive Policy Index)
print(contingency_table_pos)

# Run the chi-square test of independence (Positive Policy Index)
chi_squared_result_pos <- chisq.test(contingency_table_pos)

# Print the test result (Positive Policy Index)
print(chi_squared_result_pos)
```


# Repeat All for Negative policy

# Linear Models and Residual Analysis of RETAIL
To control for macroeconomic, socioeconomic, and political factors, we ran several linear models and used residuals to further explore relationships.

```{r linear-models, echo=TRUE}

# Run a linear model for the total positive index vs the 5 types of control_data
lm_model_neg_index <- lm(pol_indices$neg_combo_index ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity + employ_z_scores$rtrade_employment_z)

# Save the residuals of the model to regression_data$tot_pos_index_residuals
pol_indices$neg_index_residuals <- residuals(lm_model_neg_index)

# Print the summary of the linear model
summary(lm_model_neg_index)

```

# Residualized Analysis
We used residualized values to examine relationships after controlling for external factors.

```{r scatter-residual-pos-fairchance, echo=TRUE}
# Scatter plot of residualized positive policy vs. residualized fair chance job posting rate
plot_residual_neg_vs_fairchance <- ggplot(data = data.frame(y = employment_data$retail_residuals, x = pol_indices$neg_index_residuals)) +
  geom_point(aes(x = x, y = y)) +
  geom_smooth(aes(x = x, y = y), method="lm", se=FALSE)+
  labs(
    title = "Scatter Plot of Residualized Positive Policy vs. Residualized Fair Chance Job Posting Rate",
    y = "Residualized Fair Chance Job Posting Rate, Retail + Food Service",
    x = "Residualized Postitve Index"
  ) +
  theme_minimal()
if (overwrite_plots) {
  pdf("plot_residual_neg_vs_fairchance.pdf")
  print(plot_residual_neg_vs_fairchance)
  dev.off()
}
print(plot_residual_neg_vs_fairchance)
```

# Chi-Square Test of Independence
Finally, we perform chi-square tests to understand if having an above-average policy index is associated with above-average employment outcomes.

```{r chi-square-test-pos, echo=TRUE}
# Create binary variables for "above or below average" for each measure (Positive Policy Index)
policy_above_avg_neg <- ifelse(pol_indices$neg_index_residuals > 0, "Above Avg", "Below Avg")
employment_above_avg <- ifelse(employment_data$retail_residuals > 0, "Above Avg", "Below Avg")

# Create a contingency table based on the two categorical variables (Positive Policy Index)
contingency_table_neg <- table(policy_above_avg_neg, employment_above_avg)

# Print the contingency table (Positive Policy Index)
print(contingency_table_neg)

# Run the chi-square test of independence (Positive Policy Index)
chi_squared_result_neg <- chisq.test(contingency_table_pos)

# Print the test result (Positive Policy Index)
print(chi_squared_result_neg)
```