---
title: "Q2: Industry Validation Regression Models"
author: "Alayna Johnson"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=TRUE}
library(ggplot2)

# Set parameter to overwrite plots
overwrite_plots = FALSE

# Read in the data
regression_data <- read.csv("regression.csv")

employment_data <- read.csv("../Employment/Indeed/indeed-retail-foodservice-fairchance.csv")
new_employment_data <- read.csv("wide_complete_industry.csv")

control_data <- read.csv("control_indices.csv")
employment_control_data <- read.csv("employment_z_scores.csv")
pos_pol_data <- read.csv("../Policy/PositiveIndex.csv")
pos_pol_data <- pos_pol_data[1:50, ]
```


# Introduction

This analysis aims to explore the relationship between different positive policies, and fair chance hiring outcomes in retail, food service, manufacturing, home health, construction, and hospitality industries. We use linear regression models in this validation.

## Linear Models

To control for macroeconomic, socioeconomic, and political factors, we ran several linear models to explore relationships.

### Original Analysis on Fair Chance in Retail and Food Service

```{r linear-models, echo=TRUE}
# Run a linear model for fair_chance_rate vs the three columns of control_data
lm_fairchance_controls <- lm(employment_data$fair_chance_rate ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity)

# Save the residuals of the model to employment_data$fair_chance_rate_residuals
employment_data$fair_chance_rate_residuals <- residuals(lm_fairchance_controls)

# Print the summary of the linear model
summary(lm_fairchance_controls)

# Run a linear model for btb_index vs the three columns of control_data
lm_model_btb <- lm(pos_pol_data$Ban.the.Box.Index ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity)

# Save the residuals of the model to regression_data$tot_pos_index_residuals
regression_data$btb_index_residuals <- residuals(lm_model_btb)

# Print the summary of the linear model
summary(lm_model_btb)

# Run a linear model for recordclearing vs the three columns of control_data
lm_model_rec_clear <- lm(pos_pol_data$Record.Clearing.Index ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics  + control_data$diversity)

# Save the residuals of the model to regression_data$neg_combo_index_residuals
regression_data$rec_clear_index_residuals <- residuals(lm_model_rec_clear)

# Print the summary of the linear model
summary(lm_model_rec_clear)
```

### New Industry Validation Analysis

**Retail Regression**

```{r retail-validation}
# Run a linear model for avg_retail vs all control_data
lm_retail_controls <- lm(new_employment_data$avg_retail ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity + employment_control_data$rtrade_employment_z)

# Save the residuals of the model to employment_data$fair_chance_rate_residuals
retail_residuals<- residuals(lm_retail_controls)

# Print the summary of the linear model
summary(lm_retail_controls)

# Run a linear model for btb_index vs the three columns of control_data
lm_retail_btb <- lm(pos_pol_data$Ban.the.Box.Index ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity + employment_control_data$rtrade_employment_z)

# Save the residuals of the model
btb_retail_residuals <- residuals(lm_retail_btb)

# Print the summary of the linear model
summary(lm_retail_btb)

# Run a linear model for recordclearing vs the three columns of control_data
lm_retail_rec_clear <- lm(pos_pol_data$Record.Clearing.Index ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics  + control_data$diversity + employment_control_data$rtrade_employment_z)

# Save the residuals of the model to regression_data$neg_combo_index_residuals
rec_clear_retial_residuals <- residuals(lm_model_rec_clear)

# Print the summary of the linear model
summary(lm_retail_rec_clear)
```
**Food Service Regression**

```{r food-service-validation}
# Run a linear model for avg_food_service vs all control_data
lm_food_service_controls <- lm(new_employment_data$avg_food_service ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity + employment_control_data$arts_employment_z)

# Save the residuals of the model to employment_data$fair_chance_rate_residuals
food_service_residuals <- residuals(lm_food_service_controls)

# Print the summary of the linear model
summary(lm_food_service_controls)

# Run a linear model for btb_index vs the three columns of control_data
lm_food_service_btb <- lm(pos_pol_data$Ban.the.Box.Index ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity + employment_control_data$arts_employment_z)

# Save the residuals of the model
btb_food_service_residuals <- residuals(lm_food_service_btb)

# Print the summary of the linear model
summary(lm_food_service_btb)

# Run a linear model for recordclearing vs the three columns of control_data
lm_food_service_rec_clear <- lm(pos_pol_data$Record.Clearing.Index ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity + employment_control_data$arts_employment_z)

# Save the residuals of the model to regression_data$neg_combo_index_residuals
rec_clear_food_service_residuals <- residuals(lm_food_service_rec_clear)

# Print the summary of the linear model
summary(lm_food_service_rec_clear)

```
**Manufacturing Regression**

```{r manufacturing-validation}
# Run a linear model for avg_manufacturing vs all control_data
lm_manufacturing_controls <- lm(new_employment_data$avg_manufacturing ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity + employment_control_data$manu_employment_z)

# Save the residuals of the model to employment_data$fair_chance_rate_residuals
manufacturing_residuals <- residuals(lm_manufacturing_controls)

# Print the summary of the linear model
summary(lm_manufacturing_controls)

# Run a linear model for btb_index vs the three columns of control_data
lm_manufacturing_btb <- lm(pos_pol_data$Ban.the.Box.Index ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity + employment_control_data$manu_employment_z)

# Save the residuals of the model
btb_manufacturing_residuals <- residuals(lm_manufacturing_btb)

# Print the summary of the linear model
summary(lm_manufacturing_btb)

# Run a linear model for recordclearing vs the three columns of control_data
lm_manufacturing_rec_clear <- lm(pos_pol_data$Record.Clearing.Index ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity + employment_control_data$manu_employment_z)

# Save the residuals of the model to regression_data$neg_combo_index_residuals
rec_clear_manufacturing_residuals <- residuals(lm_manufacturing_rec_clear)

# Print the summary of the linear model
summary(lm_manufacturing_rec_clear)

```
**Home Health Regression**

```{r home-health-validation}
# Run a linear model for avg_home_health vs all control_data
lm_home_health_controls <- lm(new_employment_data$avg_home_health ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity + employment_control_data$edu_employment_z)

# Save the residuals of the model to employment_data$fair_chance_rate_residuals
home_health_residuals <- residuals(lm_home_health_controls)

# Print the summary of the linear model
summary(lm_home_health_controls)

# Run a linear model for btb_index vs the three columns of control_data
lm_home_health_btb <- lm(pos_pol_data$Ban.the.Box.Index ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity + employment_control_data$edu_employment_z)

# Save the residuals of the model
btb_home_health_residuals <- residuals(lm_home_health_btb)

# Print the summary of the linear model
summary(lm_home_health_btb)

# Run a linear model for recordclearing vs the three columns of control_data
lm_home_health_rec_clear <- lm(pos_pol_data$Record.Clearing.Index ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity + employment_control_data$edu_employment_z)

# Save the residuals of the model to regression_data$neg_combo_index_residuals
rec_clear_home_health_residuals <- residuals(lm_home_health_rec_clear)

# Print the summary of the linear model
summary(lm_home_health_rec_clear)

```
**Construction Regression**

```{r construction-validation}
# Run a linear model for avg_construction vs all control_data
lm_construction_controls <- lm(new_employment_data$avg_construction ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity + employment_control_data$construct_employment_z)

# Save the residuals of the model to employment_data$fair_chance_rate_residuals
construction_residuals <- residuals(lm_construction_controls)

# Print the summary of the linear model
summary(lm_construction_controls)

# Run a linear model for btb_index vs the three columns of control_data
lm_construction_btb <- lm(pos_pol_data$Ban.the.Box.Index ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity + employment_control_data$construct_employment_z)

# Save the residuals of the model
btb_construction_residuals <- residuals(lm_construction_btb)

# Print the summary of the linear model
summary(lm_construction_btb)

# Run a linear model for recordclearing vs the three columns of control_data
lm_construction_rec_clear <- lm(pos_pol_data$Record.Clearing.Index ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity + employment_control_data$construct_employment_z)

# Save the residuals of the model to regression_data$neg_combo_index_residuals
rec_clear_construction_residuals <- residuals(lm_construction_rec_clear)

# Print the summary of the linear model
summary(lm_construction_rec_clear)

```
**Hospitality Regression**

```{r hospitality-validation}
# Run a linear model for avg_hospitality vs all control_data
lm_hospitality_controls <- lm(new_employment_data$avg_hospitality ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity + employment_control_data$arts_employment_z)

# Save the residuals of the model to employment_data$fair_chance_rate_residuals
hospitality_residuals <- residuals(lm_hospitality_controls)

# Print the summary of the linear model
summary(lm_hospitality_controls)

# Run a linear model for btb_index vs the three columns of control_data
lm_hospitality_btb <- lm(pos_pol_data$Ban.the.Box.Index ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity + employment_control_data$arts_employment_z)

# Save the residuals of the model
btb_hospitality_residuals <- residuals(lm_hospitality_btb)

# Print the summary of the linear model
summary(lm_hospitality_btb)

# Run a linear model for recordclearing vs the three columns of control_data
lm_hospitality_rec_clear <- lm(pos_pol_data$Record.Clearing.Index ~ control_data$macroeconomy + control_data$socioeconomics + control_data$politics + control_data$diversity + employment_control_data$arts_employment_z)

# Save the residuals of the model to regression_data$neg_combo_index_residuals
rec_clear_hospitality_residuals <- residuals(lm_hospitality_rec_clear)

# Print the summary of the linear model
summary(lm_hospitality_rec_clear)

```



