---
title: "Q1 Research Results"
author: "Jocelyn Bliven"
date: "2024-11-07"
output: pdf_document
---

``````{r setup, include=TRUE}

library(ggplot2)

# Set parameter to overwrite plots
overwrite_plots = FALSE

# Read in data
pol_indices <- read.csv("regression.csv")
orig_employ_data <- read.csv("indeed-retail-foodservice-fairchance.csv")
new_employ_data <- read.csv("wide_complete_industry.csv")


```

## Original Plots

Linear models of positive policy index and fair chance job posting rate, negative policy index and fair chance job posting rate, and positive policy index and negative policy index

```{r linear-models, echo=TRUE}
# Run a linear model for fair_chance_rate vs the 5 types of control_data
lm_pos_vs_employ <- lm(orig_employ_data$fair_chance_rate ~ pol_indices$tot_pos_index)

# Print the summary of the linear model
summary(lm_pos_vs_employ)
```

Visualization
```{r scatter-pos-fairchance, echo=TRUE}
# Scatter plot of  positive policy vs.  fair chance job posting rate
plot_residual_pos_vs_fairchance <- ggplot(data = data.frame(y = orig_employ_data$fair_chance_rate, x = pol_indices$tot_pos_index)) +
  geom_point(aes(x = x, y = y)) +
  geom_smooth(aes(x = x, y = y), method="lm", se=FALSE)+
  labs(
    title = "Scatter Plot of Positive Policy vs. Fair Chance Job Posting Rate",
    y = "Original Fair Chance Job Posting Rate",
    x = "Postitve Index"
  ) +
  theme_minimal()
if (overwrite_plots) {
  pdf("plot_residual_pos_vs_fairchance.pdf")
  print(plot_residual_pos_vs_fairchance)
  dev.off()
}
print(plot_residual_pos_vs_fairchance)
```

Negative Policy vs Fair Chance Hiring Rate

```{r linear-models, echo=TRUE}
# Run a linear model for fair_chance_rate vs the 5 types of control_data
lm_neg_vs_employ <- lm(orig_employ_data$fair_chance_rate ~ pol_indices$neg_combo_index)

# Print the summary of the linear model
summary(lm_neg_vs_employ)
```

```{r scatter-pos-fairchance, echo=TRUE}
# Scatter plot of negative policy vs.  fair chance job posting rate
plot_residual_neg_vs_fairchance <- ggplot(data = data.frame(y = orig_employ_data$fair_chance_rate, x = pol_indices$neg_combo_index)) +
  geom_point(aes(x = x, y = y)) +
  geom_smooth(aes(x = x, y = y), method="lm", se=FALSE)+
  labs(
    title = "Scatter Plot of Negative Policy vs. Fair Chance Job Posting Rate",
    y = "Original Fair Chance Job Posting Rate",
    x = "Negative Index"
  ) +
  theme_minimal()
if (overwrite_plots) {
  pdf("plot_residual_neg_vs_fairchance.pdf")
  print(plot_residual_neg_vs_fairchance)
  dev.off()
}
print(plot_residual_neg_vs_fairchance)
```

Positive vs Negative Index

```{r linear-models, echo=TRUE}
# Run a linear model for fair_chance_rate vs the 5 types of control_data
lm_neg_vs_pos <- lm(pol_indices$tot_pos_index ~ pol_indices$neg_combo_index)

# Print the summary of the linear model
summary(lm_neg_vs_pos)
```


Visualizaiton
```{r scatter-pos-fairchance, echo=TRUE}
# Scatter plot of negative policy vs.  fair chance job posting rate
plot_residual_neg_vs_pos <- ggplot(data = data.frame(y = pol_indices$tot_pos_index, x = pol_indices$neg_combo_index)) +
  geom_point(aes(x = x, y = y)) +
  geom_smooth(aes(x = x, y = y), method="lm", se=FALSE)+
  labs(
    title = "Scatter Plot of Negative Policy vs. Positive Policy",
    y = "Positive Index",
    x = "Negative Index"
  ) +
  theme_minimal()
if (overwrite_plots) {
  pdf("plot_residual_neg_vs_pos.pdf")
  print(plot_residual_neg_vs_pos)
  dev.off()
}
print(plot_residual_neg_vs_pos)
```

## New Industry Plots

Positive 

```{r linear-models, echo=TRUE}
# Run a linear model for fair_chance_rate vs the 5 types of control_data
lm_pos_vs_retail <- lm(new_employ_data$avg_retail ~ pol_indices$tot_pos_index)

# Print the summary of the linear model
summary(lm_pos_vs_retail)

# Run a linear model for fair_chance_rate vs the 5 types of control_data
lm_pos_vs_food_service <- lm(new_employ_data$avg_food_service ~ pol_indices$tot_pos_index)

# Print the summary of the linear model
summary(lm_pos_vs_food_service)

# Run a linear model for fair_chance_rate vs the 5 types of control_data
lm_pos_vs_avg_construction <- lm(new_employ_data$avg_construction ~ pol_indices$tot_pos_index)

# Print the summary of the linear model
summary(lm_pos_vs_avg_construction)

# Run a linear model for fair_chance_rate vs the 5 types of control_data
lm_pos_vs_home_health <- lm(new_employ_data$avg_home_health ~ pol_indices$tot_pos_index)

# Print the summary of the linear model
summary(lm_pos_vs_home_health)

# Run a linear model for fair_chance_rate vs the 5 types of control_data
lm_pos_vs_manufacturing <- lm(new_employ_data$avg_manufacturing ~ pol_indices$tot_pos_index)

# Print the summary of the linear model
summary(lm_pos_vs_manufacturing)

# Run a linear model for fair_chance_rate vs the 5 types of control_data
lm_pos_vs_hospitality <- lm(new_employ_data$avg_hospitality ~ pol_indices$tot_pos_index)

# Print the summary of the linear model
summary(lm_pos_vs_hospitality)
```

Negative

```{r linear-models, echo=TRUE}
# Run a linear model for fair_chance_rate vs the 5 types of control_data
lm_neg_vs_retail <- lm(new_employ_data$avg_retail ~ pol_indices$neg_combo_index)

# Print the summary of the linear model
summary(lm_neg_vs_retail)

# Run a linear model for fair_chance_rate vs the 5 types of control_data
lm_neg_vs_food_service <- lm(new_employ_data$avg_food_service ~ pol_indices$neg_combo_index)

# Print the summary of the linear model
summary(lm_neg_vs_food_service)

# Run a linear model for fair_chance_rate vs the 5 types of control_data
lm_neg_vs_avg_construction <- lm(new_employ_data$avg_construction ~ pol_indices$neg_combo_index)

# Print the summary of the linear model
summary(lm_neg_vs_avg_construction)

# Run a linear model for fair_chance_rate vs the 5 types of control_data
lm_neg_vs_home_health <- lm(new_employ_data$avg_home_health ~ pol_indices$neg_combo_index)

# Print the summary of the linear model
summary(lm_neg_vs_home_health)

# Run a linear model for fair_chance_rate vs the 5 types of control_data
lm_neg_vs_manufacturing <- lm(new_employ_data$avg_manufacturing ~ pol_indices$neg_combo_index)

# Print the summary of the linear model
summary(lm_neg_vs_manufacturing)

# Run a linear model for fair_chance_rate vs the 5 types of control_data
lm_neg_vs_hospitality <- lm(new_employ_data$avg_hospitality ~ pol_indices$neg_combo_index)

# Print the summary of the linear model
summary(lm_neg_vs_hospitality)
```
